Detecting Treatment-Subgroup Interactions in Clustered Data with Generalized Linear Mixed-Effects Model Trees

### 2015-09-15 pkg/glmertree/paper/treatment_interactions.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@783 255f4249-2263-4417-b1ea-8498779e27a7

The algorithm initializes by setting all $b$ values to $0$, since the random-effects (and also the fixed-effects) parts are initially unknown. In every iteration, the GLM tree (i.e., the partition and corresponding fixed-effects coefficients $\beta_j$) and random-effects coefficients $b_m$ are re-estimated. The GLM tree is estimated, given the estimated $\hat{b}$ from the last iteration, and the $b$ values are estimated, given the estimated GLM tree from the current iteration. Iterations are continued until convergence, which is monitored by computing the log-likelihood criterion of the mixed-effects model in Equation~\ref{eq:mixedeffects}.

@@

Using the same data for training and evaluation of a model results in overly optimistic estimates of predictive accuracy \cite{HastyTibs09}. Therefore, GLM and GLMM trees were used for prediction of new observations from test datasets. Test datasets were generated from the same population as the training datasets. Because the cluster-specific intercepts $b_m$ were randomly generated for training as well as test datasets, test observations were from 'new' clusters. As a result, a model without random effects was used for prediction with GLMM tree.

>>>>

The algorithm initializes by setting all $b$ values to $0$, since the random-effects (and also the fixed-effects) parts are initially unknown. In every iteration, the GLM tree and random-effects coefficients $b$ are re-estimated. The GLM tree is estimated, given the estimated $\hat{b}$ from the last iteration, and the $b$ values are estimated, given the estimated GLM tree from the current iteration. Iterations are continued until convergence, which is monitored by computing the log-likelihood criterion of the mixed-effects model in Equation~\ref{eq:mixedeffects}.

@@

Using the same data for training and evaluation of a model results in overly optimistic estimates of predictive accuracy \cite{HastyTibs09}. Therefore, GLM and GLMM trees were used for prediction of new observations from test datasets. Test datasets were generated from the same population as the training datasets. Because the cluster-specific intercepts $b$ were randomly generated for training as well as test datasets, test observations were from 'new' clusters. As a result, a model without random effects was used for prediction with GLMM tree.
